---
title: "MOFA exploration"
output:
  #pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(tidyverse)
library(MOFA2)
library(here)
#library(MOFAdata)
library(ReactomePA)
library(biomaRt)
library(cowplot)
library(enrichplot)
library(dplyr)
library(clusterProfiler)
library(utils)
library(stats)
```


```{r}
# input
model <- load_model(here("models/mofa/model.hdf5"))

# gene expression data
promise_long_filtered_top <- readRDS(here::here('data/processed/expression/promise_expr_filtered_tidy_top.rds'))

# organoid morphology
umap_df <- readRDS(here::here("data/processed/morphology/umap_absolute_all_drugs_sampled.Rds"))

# organoid size
organoid_size_fit <- readRDS(here::here("data/processed/morphology/organoid_size_fit.Rds")) %>% 
  filter(!line %in% c('D055T01', 'D020T02', 'D021T01')) %>% 
  #filter(!line %in% c('D055T01','D020T02')) %>% 
  mutate(line = as.character(line)) %>% 
  dplyr::select(line, size = x, rep = replicate) %>% 
  distinct() %>% arrange(line) %>%
  mutate(line = substr(line, 1, 4)) %>% 
  mutate(rep = paste0("r", rep))

# morphology classification
organoid_morphology <- read_delim(here::here("references/imaging/visual_classification_organoids.csv"), ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::select(line = organoid, morphology = visual_inspection_v2) %>%
  mutate(line = substr(line, 1, 4)) %>% 
  mutate(morphology = if_else(is.na(morphology), "other", morphology))

# drug activity data
utils::data('aucroc', package = 'SCOPEAnalysis')
drug_activity <- aucroc %>% filter(!line %in% c('D055T01', 'D021T01', 'D054T01'))

# gene expression annotation 
intestinal_sig <- readxl::read_excel(here('data/external/expression/merloz-suarez_sigantures.xls'), 
                                     sheet = 1, skip = 4) %>% .[,1:4] %>%
  gather(signature, symbol) %>% drop_na() %>% 
  mutate(symbol = gsub('\\*', '', symbol))

cris_sig <- readxl::read_excel(here('data/external/expression/cris/41467_2017_BFncomms15107_MOESM422_ESM.xlsx'), sheet = 1, skip = 2) %>% 
  dplyr::rename(symbol = `Gene ID`, signature = `CRIS Class`)

# organoid mutation
mofa_genetics <- read_delim(here::here("data/processed/mutation/Table-S2_Mutations_PDOs_RevisionV4.csv"), delim = ";") %>% 
    janitor::clean_names() %>% 
    mutate(sample = substr(sample, 2,nchar(sample)-1)) %>% 
    mutate(sample = paste0("D", sample)) %>% 
    dplyr::filter(!sample %in% c("D021", "D015")) %>%
    expand_grid(replicate = c("r1", "r2")) %>% 
    mutate(sample = paste0(sample, "_", replicate)) %>% 
    dplyr::select(sample, feature = symbol, everything()) %>% dplyr::select(sample, feature) %>% 
    mutate(value = 1) %>%
    complete(sample, feature, fill = list(value = 0)) %>% 
    distinct(sample, feature, value) %>%
    mutate(view = "mutation")


weights <- model@expectations$Z$single_group %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()
loadings_size <- model@expectations$W$size_view %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()
loadings_expression <- model@expectations$W$expression %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()
loadings_morphology <- model@expectations$W$morphology %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()
loadings_drug_activity <- model@expectations$W$drug_activity %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()
loadings_mutation <- model@expectations$W$mutation %>% as.data.frame() %>% rownames_to_column("id") %>% as_tibble() %>% janitor::clean_names()


```

# QC

```{r}
plot_data_overview(model) + 
  ggsave(here::here("reports/figures/mofa/data_overview.pdf"))
```

```{r}
df <- head(model@cache$variance_explained$r2_total[[1]]) %>%
  as.data.frame() %>%
  rownames_to_column("feature") %>%
  magrittr::set_colnames(c("feature", "var")) %>%
  mutate(var = var/100)

gg_ve_feature <- df %>%
  ggplot(aes(feature, var)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  theme_cowplot() + 
  labs(y = "R2")

gg_ve_feature + 
  ggsave(here::here("reports/figures/mofa/var_explained_feature.pdf"))
```


Breakdown by factors 

```{r}
model@cache$variance_explained$r2_per_factor$single_group
```

```{r}
model@cache$variance_explained$r2_per_factor$single_group %>% as.data.frame() %>% 
  rownames_to_column("factor") %>% 
  mutate(overall_variance = expression + morphology_view + size_view + drug_activity + mutation) %>%
  mutate(overall_variance = overall_variance/5) %>%
  ggplot(aes(factor, overall_variance)) + 
  geom_point(size = 2) + 
  theme_cowplot() + 
  ggsave(here::here("reports/figures/mofa/var_explained_avg.pdf"))
```

```{r}
# variance explained for different factors
model@cache$variance_explained$r2_per_factor$single_group %>% as.data.frame() %>% 
    rownames_to_column("factor") %>% 
    mutate(overall_variance = expression + morphology_view + size_view + drug_activity) %>% 
    filter(factor != "Factor3") %>% dplyr::select(-factor) %>% colSums()
```


```{r}
gg_ve <- plot_variance_explained(model, x="view", y="factor") + coord_flip() +
  theme_cowplot() 
gg_ve  + 
  #coord_equal() +
  ggsave(here::here("reports/figures/mofa/var_explained.pdf"))
  
```

```{r}
ph_cor <- weights %>% 
  as.data.frame() %>% 
  column_to_rownames("id") %>% 
  cor() %>% 
  pheatmap::pheatmap()
```

```{r}
weights %>% 
  as.data.frame() %>% 
  column_to_rownames("id") %>% 
  cor() %>% as.vector() %>% 
  .[. != 1] %>% 
  summary()
```


```{r}
plot_factor(model, 
  factor = 1:3
)
```

# factor overview

```{r}
umap_factor <- weights %>% 
  separate(id, c("line", "replicate"), sep = "_") %>% 
  mutate(line = paste0(line, "T01"),
         replicate = substr(replicate, 2,2)) %>% 
  left_join(umap_df %>% 
  filter(drug == "DMSO"), .) %>% 
  filter(line != "D020T02") 

gg_factor_umap <- umap_factor %>% 
  dplyr::select(-size_factor) %>%
  pivot_longer(cols = contains("factor"), names_to = "number", values_to = "value") %>%
  ggplot(aes(v1, v2, color = value)) + 
  ggrastr::geom_point_rast(alpha = 0.1, size = 0.35) + 
  scale_color_viridis_c() +
  cowplot::theme_cowplot() +
  labs(x = "UMAP 1",
       y = "UMAP 2",
       color = "factor value") + 
  facet_wrap(~ number) +
  theme(legend.position = "bottom") + 
    coord_fixed()

gg_factor_umap + ggsave(here("reports/figures/mofa/factor_overview.pdf"), height = 6 , width = 6)
```



```{r}
gg_f12 <- weights %>% 
  dplyr::rename(line = id) %>%
  left_join(organoid_size_fit %>% mutate(line = paste0(line, "_", rep))) %>%
  left_join(organoid_morphology %>% 
            mutate(line = substr(line, 1, 4)) %>% 
  expand_grid(., rep = c("r1", "r2")) %>% 
  mutate(sample = paste0(line, "_", rep)) %>% dplyr::select(-line, line = sample)) %>%
  distinct() %>%
  ggplot(aes(factor1, factor2, label = line, color = size, shape = morphology)) + 
  geom_point(size = 4) + 
  ggrepel::geom_text_repel(color = "black") + 
  scale_color_viridis_c() + 
  theme_cowplot() + 
  coord_fixed()

gg_f12 + ggsave(here("reports/figures/mofa/factor_12_plot.pdf"), height = 6 , width = 6)
```


```{r}
gg_f23 <- weights %>% 
  dplyr::rename(line = id) %>%
  left_join(organoid_size_fit %>% mutate(line = paste0(line, "_", rep))) %>%
  ggplot(aes(factor1, factor3, label = line)) + 
  geom_point(size = 2) + 
  ggrepel::geom_text_repel(color = "black") + 
  scale_color_viridis_c() + 
  theme_cowplot() + 
    theme_cowplot() + 
  coord_fixed() 

gg_f23 + ggsave(here("reports/figures/mofa/factor_23_plot.pdf"), height = 5 , width = 5)
```

```{r}
cowplot::plot_grid(plot_grid(gg_ve, gg_ve_feature, gg_f12, ncol = 3),
                   gg_factor_umap,
                   ncol = 1
                   ) + 
  ggsave(here("reports/figures/mofa/mofa_overview.pdf"), height = 10 , width = 15)
```


## mutations

```{r}
loadings_mutation %>%
  column_to_rownames("id") %>% 
  pheatmap::pheatmap()
```

```{r}
df <- loadings_mutation %>% 
  gather(key = "factor", value = "weight", -id) 

gg_weight_mutation <- df %>%
  ggplot(aes(factor, weight, label = id)) +
  geom_point() + 
  theme_cowplot() + 
  ggrepel::geom_text_repel(data = df %>% filter(id %in% c("NRAS", "ERBB2", "PTEN", "KRAS", "PIK3CA"))) + 
  geom_hline(yintercept = 0, linetype = "dashed")

gg_weight_mutation + ggsave(here::here("reports/figures/mofa/mutation_weight.pdf"), width = 5, height = 5)
```



## drug response


```{r}
## load drug annotation
drug_anno <- readxl::read_excel(here::here('references/layouts/Compound_Annotation_Libraries_New.xlsx')) %>% distinct(drug = `Compound name`, target = `Primary Target`)

loadings_drug_activity_tidy <- loadings_drug_activity %>% 
  mutate(drug = substr(id, 7, nchar(.))) %>% 
  dplyr::select(-id) %>% 
  arrange(desc(factor1)) %>% 
  left_join(drug_anno)


# 
# top_n <- 20
# 
# top_drugs <- rbind(
#   loadings_drug_activity_tidy %>% arrange(factor1) %>% head(top_n),
#   loadings_drug_activity_tidy %>% arrange(factor1) %>% tail(top_n),
#   loadings_drug_activity_tidy %>% arrange(factor2) %>% head(top_n),
#   loadings_drug_activity_tidy %>% arrange(factor2) %>% tail(top_n),
#   loadings_drug_activity_tidy %>% arrange(factor3) %>% head(top_n),
#   loadings_drug_activity_tidy %>% arrange(factor3) %>% tail(top_n)
# ) %>% distinct()
# 
# top_drugs %>% 
#   column_to_rownames("drug") %>% 
#   pheatmap::pheatmap()
```

```{r}
df <- loadings_drug_activity_tidy %>% dplyr::select(-target) %>% 
  gather(factor_name, value, -drug) %>% 
  distinct() 

row_anno <-  df %>% left_join(loadings_drug_activity_tidy) %>% dplyr::select(drug, target) %>%
  mutate(id = paste0(drug, "_", target)) %>% distinct() %>% as.data.frame() %>% column_to_rownames("id") 
           

df %>%
  spread(drug, value) %>% 
  dplyr::select(-factor_name) %>%
  cor() %>% 
  as.matrix() %>% 
  pheatmap::pheatmap(show_rownames = FALSE, 
                     show_colnames = FALSE)
```




```{r}
drug_activity_test <- loadings_drug_activity_tidy %>% semi_join(loadings_drug_activity_tidy %>% 
                                            dplyr::count(target) %>% 
                                            filter(n >=5)) 
```


```{r}
drug_activity_test_result <- rbind(
  lm(factor1 ~ target, data = drug_activity_test) %>% summary() %>% broom::tidy() %>% mutate(factor = "factor1"),
  lm(factor2 ~ target, data = drug_activity_test) %>% summary() %>% broom::tidy() %>% mutate(factor = "factor2"),
  lm(factor3 ~ target, data = drug_activity_test) %>% summary() %>% broom::tidy() %>% mutate(factor = "factor3")) 

df <- drug_activity_test_result %>% 
  filter(term != "(Intercept)") %>% 
  dplyr::select(term, statistic, factor) %>% 
  mutate(term = substr(term, 7, nchar(term)))

df %>% 
  spread(factor, statistic) %>% 
  column_to_rownames("term") %>% 
  pheatmap::pheatmap()
```

```{r}
drug_activity_test_result %>% 
  mutate(fdr = p.adjust(p.value, method = "BH")) %>% 
  filter(fdr <= 0.2) %>% 
  filter(term != "(Intercept)") %>% 
  arrange(factor, statistic)
```



## exporting factors loadings for enrichment

```{r}
# top 100 genes
loadings_expression %>% arrange(desc(factor1)) %>% head(100) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  arrange(desc(id)) %>%
  dplyr::select(id) %>% write_csv(here::here("reports/tables/factor1_top.csv"))

loadings_expression %>% arrange(desc(factor2)) %>% head(100) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  dplyr::select(id) %>% write_csv(here::here("reports/tables/factor2_top.csv"))

loadings_expression %>% arrange(desc(factor3)) %>% head(100) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  dplyr::select(id) %>% write_csv(here::here("reports/tables/factor3_top.csv"))

# all genes
loadings_expression %>% arrange(desc(factor1)) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  arrange(desc(id)) %>%
  dplyr::select(id, factor1) %>% write_csv(here::here("reports/tables/factor1_all.csv"))

loadings_expression %>% arrange(desc(factor2)) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  dplyr::select(id, factor2) %>% write_csv(here::here("reports/tables/factor2_all.csv"))

loadings_expression %>% arrange(desc(factor3)) %>% 
  dplyr::mutate(id = substr(id, 1, nchar(id)-11)) %>% 
  dplyr::select(id, factor3) %>% write_csv(here::here("reports/tables/factor3_all.csv"))
```

# factor 1 

## size 

```{r}
gg_sizef1 <- weights %>% 
  dplyr::rename(line = id) %>%
  left_join(organoid_size_fit %>% mutate(line = paste0(line, "_", rep))) %>%
  left_join(organoid_morphology %>% 
            mutate(line = substr(line, 1, 4)) %>% 
  expand_grid(., rep = c("r1", "r2")) %>% 
  mutate(sample = paste0(line, "_", rep)) %>% dplyr::select(-line, line = sample)) %>%
  distinct() %>%
  ggplot(aes(factor1, size, label = line)) + 
  geom_smooth(method = "lm", se = FALSE, color = "grey") +
  geom_point(size = 2) + 
  ggrepel::geom_text_repel(color = "black") + 
  scale_color_viridis_c() + 
  labs(y = "expected size") +
  theme_cowplot() + 
  coord_fixed(ratio = 2)

gg_sizef1 + ggsave(here::here("reports/figures/mofa/f1_size.pdf"), width = 5, height = 5)
```





## gene expression

```{r}
gg_f1_geneexp <- plot_weights(model,
  view = "expression",
  factor = 1,
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
) + 
  theme_cowplot() + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y =element_blank()) 

gg_f1_geneexp + 
  ggsave(here::here("reports/figures/mofa/f1_genexp.pdf"), width = 4, height = 4)
```
growth control via IGF signaling
cell adhesion via FYN signaling, high fibronectin expression
DLX- TGFb and BMP inhibiting transcription factor

```{r}
df <- loadings_expression %>% 
  separate(id, c("symbol"), sep = "_exp") %>% 
  filter(symbol != "") %>%
  left_join(promise_long_filtered_top %>% dplyr::select(symbol, entrez) %>% distinct()) %>%
  arrange(desc(factor1)) %>%
  mutate(order = nrow(.):1)


```



```{r}
df <- loadings_expression %>% 
  separate(id, c("symbol"), sep = "_exp") %>% 
  filter(symbol != "") %>%
  left_join(promise_long_filtered_top %>% dplyr::select(symbol, entrez) %>% distinct()) %>%
  arrange(desc(factor1))
  

ranks_1 <- setNames(df$factor1, as.character(df$entrez))
ranks_1 <- sort(ranks_1, decreasing = T)

## Reactome enrichment analysis
gse_reactome_1 <- gsePathway(
  geneList = ranks_1,
  organism = 'human',
  #nPerm = 1e5,
  #minGSSize = 100,
  #maxGSSize = 500,
  pvalueCutoff = 0.2
)

reactome <- pairwise_termsim(gse_reactome_1) 
gg_f1_emap <- emapplot(reactome, color = "NES",
         cex_label_category = .8,
         cex_line = 1) + 
  scale_color_viridis_c() + 
  labs(color = "NES")

gg_f1_emap + 
  ggsave(here::here("reports/figures/mofa/f1_emap.pdf"), width = 4, height = 4)
```


```{r, eval = FALSE}
gse_go <- gseGO(
  geneList = ranks_1,
  OrgDb = org.Hs.eg.db,
  ont = 'BP',
  # nPerm = 1e5,
  # minGSSize = 10,
  # maxGSSize = 500,
  pvalueCutoff = 0.2
)

go <- pairwise_termsim(gse_go) 
emapplot(go, color = "NES")
```
### suarez signature

```{r}
# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor1, symbol) %>% distinct() %>%
  arrange(desc(factor1))
ranks_symbol = setNames(df$factor1, as.character(df$symbol))

gse_sig <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = intestinal_sig,
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_sig_tbl <- as_tibble(gse_sig)
```


```{r}
## proliferation
gg_f1prolif <- gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[1],
  title = paste0(gse_sig$ID[1], 
                ' (p = ', round(gse_sig_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_sig_tbl$NES[1], 2), ')')
)

gg_f1prolif + 
  ggsave(here::here("reports/figures/mofa/f1_prolif.pdf"), width = 4, height = 3)
```



```{r}
## late TA signature
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[2], 
  title = paste0(gse_sig$ID[2], 
                ' (p = ', round(gse_sig_tbl$pvalue[2], 3), 
                '; NES = ', round(gse_sig_tbl$NES[2], 2), ')')
) 
```


```{r}
## LGR5
gg_f1lgr <- gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[3],
  title = paste0(gse_sig$ID[3], 
                ' (p = ', round(gse_sig_tbl$pvalue[3], 3), 
                '; NES = ', round(gse_sig_tbl$NES[3], 2), ')')
)

gg_f1lgr + 
  ggsave(here::here("reports/figures/mofa/f1_lgr.pdf"), width = 4, height = 3)
```



```{r}
## EPH
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[4],
  title = paste0(gse_sig$ID[4], 
                ' (p = ', round(gse_sig_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_sig_tbl$NES[1], 2), ')')
)
```

### CRIS signature

```{r}
# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor1, symbol) %>% distinct() %>%
  arrange(desc(factor1))
ranks_symbol = setNames(df$factor1, as.character(df$symbol))

gse_cris <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = cris_sig %>% dplyr::select(signature, symbol),
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_cris_tbl <- as_tibble(gse_cris)

gse_cris_tbl
```

```{r}
## proliferation
gg_f1_crisc <- gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[1],
  title = paste0(gse_cris$ID[1], 
                ' (p = ', round(gse_cris_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_cris_tbl$NES[1], 2), ')')
)

gg_f1_crisc + 
  ggsave(here::here("reports/figures/mofa/f1_crisc.pdf"), width = 4, height = 3)
```

```{r}
gg_f1_crisd <- gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[2],
  title = paste0(gse_cris$ID[2], 
                ' (p = ', round(gse_cris_tbl$pvalue[2], 3), 
                '; NES = ', round(gse_cris_tbl$NES[2], 2), ')')
)

gg_f1_crisd + 
  ggsave(here::here("reports/figures/mofa/f1_crisd.pdf"), width = 4, height = 3)
```


## drug sensitivity

```{r}

aggregate_drugs <- drug_activity_test %>% group_by(target) %>% summarise_at(vars(contains("factor")), funs("median"))

include_drugs <- aggregate_drugs %>% arrange(desc(factor1)) %>% .$target

gg_f1_drug <- drug_activity_test %>%
  drop_na() %>%
  mutate(target = factor(target, levels = include_drugs)) %>%
  ggplot(aes(y = target, x = factor1)) + 
  #geom_point() + 
  geom_vline(xintercept = 0, color = "grey") + 
  ggridges::geom_density_ridges() +
  
  #coord_flip() + 
  cowplot::theme_cowplot()

gg_f1_drug + ggsave(here::here("reports/figures/mofa/f1_ridge_drug.pdf"), width = 4, height = 3)
```

```{r}

```


median weighting within the factor was strongest for activity of MEK, IGF1R inhibitors. mTOR inhibitors ranked among the strongest drugs as well. In contrast, EGFR inhibitor activity had the most negative median contribution to the factor. Put differently, organoid lines that had a high score for factor 1 tended to be sensitive to MEK and IGF-1R inhibitors and less responsive to EGFR and Syk inhibitors. 


EGFR inhibitor activity has a significant negative contribution to the factor. This means that lines with a high factor1 score, show little to no activity when treated with EGFR inhibitors.


```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

no_egfr <- c("Tyrphostin 9" # IC50 at >100uM, binds PDGFR
             )

set.seed(123)
doi <- drug_activity_test %>% 
  dplyr::filter(target == "EGFR") %>% 
  dplyr::filter(drug != no_egfr) %>%
  sample_n(9) %>%
  .$drug 


gg_f1_egfr <- drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor1, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~ drug) + 
  coord_fixed(ratio = 5) + 
  scale_y_continuous(limits = c(0.5, 1))


gg_f1_egfr + ggsave(here::here("reports/figures/mofa/f1_egfr.pdf"), width = 6, height = 6)

```

```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

set.seed(123)
doi <- drug_activity_test %>% 
  dplyr::filter(target == "MEK") %>% 
   sample_n(9) %>%
  .$drug 


gg_f1_mek <- drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor1, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ drug) + 
  coord_fixed(ratio = 5) + 
  scale_y_continuous(limits = c(0.5, 1))

gg_f1_mek + ggsave(here::here("reports/figures/mofa/f1_mek.pdf"), width = 6, height = 6)
```

```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

set.seed(123)
doi <- drug_activity_test %>% 
  dplyr::filter(target == "IGF-1R") %>% 
  .$drug


gg_f1_igf <- drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor1, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  facet_wrap(~ drug) + 
  coord_fixed(ratio = 5) + 
  scale_y_continuous(limits = c(0.5, 1))

gg_f1_igf + ggsave(here::here("reports/figures/mofa/f1_igf.pdf"), width = 6, height = 6)
```


```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

set.seed(123)
doi <- drug_activity_test %>% 
  dplyr::filter(target == "mTOR") %>% 
  sample_n(9, replace = TRUE) %>%
  .$drug 


gg_f1_mtor <- drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor1, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "green") + 
  facet_wrap(~ drug) + 
  coord_fixed(ratio = 5) + 
  scale_y_continuous(limits = c(0.5, 1))

gg_f1_mtor + ggsave(here::here("reports/figures/mofa/f1_mtor.pdf"), width = 6, height = 6)
```
within the group of mTOR inhibitors, activity towards treatment with the small molecule WYE-132 had the strongest contribution to the factor. WYE-132 is an ATP competitive inhibitor of mTORC1 and mTORC2

### pick of drug sensitivity

```{r}
gg_f1_egfr_pick <- drug_activity_joined %>% 
  filter(drug %in% c("OSI-420", "Trametinib (GSK1120212)", "BMS-536924")) %>% 
  ggplot(aes(factor1, auroc, color = drug)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, aes(group = drug)) + 
  coord_equal(ratio = 10) + 
  theme(legend.position = "bottom")
  

gg_f1_egfr_pick + ggsave(here::here("reports/figures/mofa/f1_drug_pick.pdf"), width = 4, height = 4)
```


Next we wondered how treatment with members from these highly active groups would change the phenotype of organoids 

## mutation

```{r}
df <- mofa_genetics %>% 
  mutate(sample = substr(sample, 1, 4)) %>% 
  distinct() %>%
  filter(feature %in% c("NRAS"))

drug_activity_joined %>%
  dplyr::select(-line, sample = id) %>% 
  filter(drug %in% c("Selumetinib (AZD6244)")) %>%
  left_join(df) %>% 
  drop_na() %>%
  mutate(value = if_else(value == 0, "WT", "mut")) %>%
  ggplot(aes(value, auroc)) + 
  geom_point() + 
  ggsignif::geom_signif(comparisons = list(c("WT", "mut"))) + 
  theme_cowplot() + 
  facet_grid(drug ~ feature)
  
```

## panel

```{r, eval = FALSE}
plot_grid(plot_grid(gg_sizef1, gg_f1_geneexp, align = "hv"), 
          gg_f1_emap,
          plot_grid(gg_f1prolif, gg_f1_crisc, gg_f1_crisd, ncol = 1),
          gg_f1_drug
          )



```


# factor 2

## morphology

```{r}
gg_cystic <- weights %>% 
  dplyr::rename(line = id) %>%
  left_join(organoid_morphology %>% 
              mutate(line = substr(line, 1, 4)) %>% 
    expand_grid(., rep = c("r1", "r2")) %>% 
      mutate(sample = paste0(line, "_", rep)) %>% dplyr::select(-line, line = sample)) %>%
  distinct() %>%
  ggplot(aes(factor1, factor2, label = line, color = morphology)) + 
  geom_point(size = 4) + 
  ggrepel::geom_text_repel(color = "black") + 
  theme_cowplot() + 
  scale_color_brewer(type = "qual")


gg_cystic + ggsave(here::here("reports/figures/mofa/f2_morph.pdf"), width = 6, height = 6)
```

## gene expression


```{r}
gg_f2_geneexp <- plot_weights(model,
  view = "expression",
  factor = 2,
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
) + 
  theme_cowplot() + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y =element_blank()) 

gg_f2_geneexp + 
  ggsave(here::here("reports/figures/mofa/f2_genexp.pdf"), width = 4, height = 4)
```
growth control via IGF signaling
cell adhesion via FYN signaling, high fibronectin expression
DLX- TGFb and BMP inhibiting transcription factor


### reactome and GO

```{r}
df <- loadings_expression %>% 
  separate(id, c("symbol"), sep = "_exp") %>% 
  filter(symbol != "") %>%
  left_join(promise_long_filtered_top %>% dplyr::select(symbol, entrez) %>% distinct()) %>%
  arrange(desc(factor2))
  

ranks_2 <- setNames(df$factor2, as.character(df$entrez))
ranks_2 <- sort(ranks_2, decreasing = T)

## Reactome enrichment analysis
gse_reactome_2 <- gsePathway(
  geneList = ranks_2,
  organism = 'human',
  #nPerm = 1e5,
  #minGSSize = 100,
  #maxGSSize = 500,
  pvalueCutoff = 0.2
)

reactome <- pairwise_termsim(gse_reactome_2) 

gg_f2_emap <- emapplot(reactome, color = "NES",
         cex_label_category = .8,
         cex_line = 1) + 
  scale_color_viridis_c() + 
  labs(color = "NES")

gg_f2_emap + 
  ggsave(here::here("reports/figures/mofa/f2_emap.pdf"), width = 4, height = 4)

```


```{r, eval = FALSE}
gse_go <- gseGO(
  geneList = ranks_2,
  OrgDb = org.Hs.eg.db,
  ont = 'BP',
  # nPerm = 1e5,
  # minGSSize = 10,
  # maxGSSize = 500,
  pvalueCutoff = 0.2
)

go <- pairwise_termsim(gse_go) 
emapplot(go, color = "NES")
```

### suarez signature

```{r}
# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor2, symbol) %>% distinct() %>%
  arrange(desc(factor2))
ranks_symbol = setNames(df$factor2, as.character(df$symbol))

gse_sig <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = intestinal_sig,
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_sig_tbl <- as_tibble(gse_sig)
```


```{r}
## lgr5
gg_f2lgr5 <- gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[1],
  title = paste0(gse_sig$ID[1], 
                ' (p = ', round(gse_sig_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_sig_tbl$NES[1], 2), ')')
)

gg_f2lgr5 + 
  ggsave(here::here("reports/figures/mofa/f2_lgr5.pdf"), width = 4, height = 3)
```


```{r}
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[2], 
  title = paste0(gse_sig$ID[2], 
                ' (p = ', round(gse_sig_tbl$pvalue[2], 3), 
                '; NES = ', round(gse_sig_tbl$NES[2], 2), ')')
) 
```


```{r}
## proliferation
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[3],
  title = paste0(gse_sig$ID[3], 
                ' (p = ', round(gse_sig_tbl$pvalue[3], 3), 
                '; NES = ', round(gse_sig_tbl$NES[3], 2), ')')
)
```


```{r}
## proliferation
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[4],
  title = paste0(gse_sig$ID[4], 
                ' (p = ', round(gse_sig_tbl$pvalue[4], 3), 
                '; NES = ', round(gse_sig_tbl$NES[4], 2), ')')
)
```

### CRIS signature

```{r}
# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor2, symbol) %>% distinct() %>%
  arrange(desc(factor2))
ranks_symbol = setNames(df$factor2, as.character(df$symbol))

gse_cris <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = cris_sig %>% dplyr::select(signature, symbol),
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_cris_tbl <- as_tibble(gse_cris)

gse_cris_tbl
```

```{r}
## proliferation
gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[3],
  title = paste0(gse_cris$ID[3], 
                ' (p = ', round(gse_cris_tbl$pvalue[3], 5), 
                '; NES = ', round(gse_cris_tbl$NES[3], 2), ')')
)
```


```{r}
## proliferation
gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[1],
  title = paste0(gse_cris$ID[1], 
                ' (p = ', round(gse_cris_tbl$pvalue[1], 5), 
                '; NES = ', round(gse_cris_tbl$NES[1], 2), ')')
)

```

```{r}
gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[4],
  title = paste0(gse_cris$ID[4], 
                ' (p = ', round(gse_cris_tbl$pvalue[4], 5), 
                '; NES = ', round(gse_cris_tbl$NES[4], 2), ')')
)
```


## drug sensitivity

```{r}

aggregate_drugs <- drug_activity_test %>% group_by(target) %>% summarise_at(vars(contains("factor")), funs("median"))

include_drugs <- aggregate_drugs %>% arrange(desc(factor2)) %>% .$target

gg_f2_drug <- drug_activity_test %>%
  drop_na() %>%
  mutate(target = factor(target, levels = include_drugs)) %>%
  ggplot(aes(y = target, x = factor2)) + 
  geom_vline(xintercept = 0, color = "grey") +
  #geom_point() + 
  ggridges::geom_density_ridges() +
   
  #coord_flip() + 
  cowplot::theme_cowplot()

gg_f2_drug + ggsave(here::here("reports/figures/mofa/f2_ridge_drug.pdf"), width = 4, height = 3)
```

median weighting within the factor was strongest for activity of Wnt/bcatenin, Src, EGFR and FAK inhibitors. In contrast, MEK, ERK inhibitor activity had among the most negative median contribution to the factor, similarly mTOR inhibitors, albeit not statistically significant. Put differently, organoid lines that had a high score for factor 2 tended to be sensitive to Wnt and EGFR targeting and less responsive to MEK and ERK inhibitors. 



```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

doi <- drug_activity_test %>% 
  dplyr::filter(target == "EGFR") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor2, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~ drug)
```

```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

doi <- drug_activity_test %>% 
  dplyr::filter(target == "MEK") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor2, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ drug)
```

```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

doi <- drug_activity_test %>% 
  dplyr::filter(target == "ERK") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor2, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ drug)
```


```{r}
doi <- drug_activity_test %>% 
  dplyr::filter(target == "Wnt/beta-catenin") %>% 
  .$drug 


gg_f2_wnt <- drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  filter(drug != "WIKI4") %>%
  filter(drug != "Wnt agonist 1") %>%
  ggplot(aes(factor2, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  facet_wrap(~ drug)

gg_f2_wnt + ggsave(here::here("reports/figures/mofa/f2_wnt.pdf"), width = 6, height = 6)
```




```{r}
drug_activity_joined <- weights %>%
  mutate(id = substr(id, 1, 4)) %>% 
  group_by(id) %>% 
 summarise_all(funs(mean)) %>% 
  mutate(line = paste0(id, "T01")) %>%
  left_join(drug_activity)

doi <- drug_activity_test %>% 
  dplyr::filter(target == "mTOR") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor2, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "green") + 
  facet_wrap(~ drug)
```
again, within the group of mTOR inhibitors, activity towards treatment with the small molecule WYE-132 had the strongest contribution to the factor - showing less activity in factor 2 high organoids compared to factor 2 low organoid lines.

### pick of drug sensitivity

```{r}
gg_f2_pick <- drug_activity_joined %>% 
  filter(drug %in% c("PRI-724", "Trametinib (GSK1120212)", "Ulixertinib (BVD-523, VRT752271)")) %>% 
  ggplot(aes(factor2, auroc, color = drug)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, aes(group = drug)) + 
  coord_equal(ratio = 10) + 
  theme(legend.position = "bottom")
  

gg_f2_pick + ggsave(here::here("reports/figures/mofa/f2_drug_pick.pdf"), width = 4, height = 4)
```

# factor 3

## morphology

## gene expression

```{r}
plot_weights(model,
  view = "expression",
  factor = 3,
  nfeatures = 10,     # Number of features to highlight
  scale = T,          # Scale weights from -1 to 1
  abs = F             # Take the absolute value?
)
```

NRX cell adhesion
xenobiotic metabolism

### reactome and GO

```{r}
df <- loadings_expression %>% 
  separate(id, c("symbol"), sep = "_exp") %>% 
  filter(symbol != "") %>%
  left_join(promise_long_filtered_top %>% dplyr::select(symbol, entrez) %>% distinct()) %>%
  arrange(desc(factor3))
  

ranks_3 <- setNames(df$factor3, as.character(df$entrez))
ranks_3 <- sort(ranks_3, decreasing = T)

## Reactome enrichment analysis
gse_reactome_3 <- gsePathway(
  geneList = ranks_3,
  organism = 'human',
  #nPerm = 1e5,
  #minGSSize = 100,
  #maxGSSize = 500,
  pvalueCutoff = 0.2
)

reactome <- pairwise_termsim(gse_reactome_3) 
emapplot(reactome, color = "NES")
```


```{r, eval = FALSE}
gse_go <- gseGO(
  geneList = ranks_3,
  OrgDb = org.Hs.eg.db,
  ont = 'BP',
  # nPerm = 1e5,
  # minGSSize = 10,
  # maxGSSize = 500,
  pvalueCutoff = 0.2
)

go <- pairwise_termsim(gse_go) 
emapplot(go, color = "NES")
```

### suarez signature

```{r}
# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor3, symbol) %>% distinct() %>%
  arrange(desc(factor3))
ranks_symbol = setNames(df$factor3, as.character(df$symbol))

gse_sig <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = intestinal_sig,
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_sig_tbl <- as_tibble(gse_sig)
```

```{r}
## proliferation
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[1],
  title = paste0(gse_sig$ID[1], 
                ' (p = ', round(gse_sig_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_sig_tbl$NES[1], 2), ')')
)
```


```{r}
## lgr5 signature
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[2], 
  title = paste0(gse_sig$ID[2], 
                ' (p = ', round(gse_sig_tbl$pvalue[2], 3), 
                '; NES = ', round(gse_sig_tbl$NES[2], 2), ')')
) 
```


```{r}
## proliferation
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[3],
  title = paste0(gse_sig$ID[3], 
                ' (p = ', round(gse_sig_tbl$pvalue[3], 3), 
                '; NES = ', round(gse_sig_tbl$NES[3], 2), ')')
)
```


```{r}
## proliferation
gseaplot2(
  gse_sig, geneSetID = gse_sig$ID[4],
  title = paste0(gse_sig$ID[4], 
                ' (p = ', round(gse_sig_tbl$pvalue[4], 3), 
                '; NES = ', round(gse_sig_tbl$NES[4], 2), ')')
)
```

### CRIS signature

```{r}
set.seed(1234)

# run gsea with clusterprofiler
df = loadings_expression %>% drop_na() %>%
  mutate(symbol = substr(id, 1, nchar(id)-11)) %>%
  dplyr::select(-id) %>%
  left_join(promise_long_filtered_top) %>%
  dplyr::select(factor3, symbol) %>% distinct() %>%
  arrange(desc(factor3))
ranks_symbol = setNames(df$factor3, as.character(df$symbol))

gse_cris <- GSEA(
  geneList = ranks_symbol,
  TERM2GENE = cris_sig %>% dplyr::select(signature, symbol),
  nPerm = 1e5,
  minGSSize = 1,
  maxGSSize = 1000,
  pvalueCutoff = 1
)

## output as tibble
gse_cris_tbl <- as_tibble(gse_cris)

gse_cris_tbl
```

```{r}
## proliferation
gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[1],
  title = paste0(gse_cris$ID[1], 
                ' (p = ', round(gse_cris_tbl$pvalue[1], 3), 
                '; NES = ', round(gse_cris_tbl$NES[1], 2), ')')
)

```

```{r}
gseaplot2(
  gse_cris, geneSetID = gse_cris$ID[2],
  title = paste0(gse_cris$ID[2], 
                ' (p = ', round(gse_cris_tbl$pvalue[2], 3), 
                '; NES = ', round(gse_cris_tbl$NES[2], 2), ')')
)
```


## drug sensitivity

```{r}

aggregate_drugs <- drug_activity_test %>% group_by(target) %>% summarise_at(vars(contains("factor")), funs("median"))

include_drugs <- aggregate_drugs %>% arrange(desc(factor3)) %>% .$target

drug_activity_test %>%
  drop_na() %>%
  mutate(target = factor(target, levels = include_drugs)) %>%
  ggplot(aes(y = target, x = factor3)) + 
  geom_vline(xintercept = 0) + 
  #geom_point() + 
  ggridges::geom_density_ridges() +
  
  #coord_flip() + 
  cowplot::theme_cowplot()
```

median weighting within the factor was strongest for activity of AURK inhibitors. In contrast, GSK and EGFR inhibitor activity had the most negative median contribution to the factor. Put differently, organoid lines that had a high score for factor 3 tended to be sensitive to ARK targeting and less responsive to GSK and EGFR inhibitors. 



```{r}
doi <- drug_activity_test %>% 
  dplyr::filter(target == "EGFR") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor3, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~ drug)
```

```{r}
doi <- drug_activity_test %>% 
  dplyr::filter(target == "GSK-3") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor3, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ drug)
```

```{r}
doi <- drug_activity_test %>% 
  dplyr::filter(target == "Aurora Kinase") %>% 
  .$drug 


drug_activity_joined %>% 
  filter(drug %in% doi) %>% 
  ggplot(aes(factor3, auroc)) +
  geom_point() + 
  cowplot::theme_cowplot() + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  facet_wrap(~ drug)
```



again, within the group of mTOR inhibitors, activity towards treatment with the small molecule WYE-132 had the strongest contribution to the factor - showing less activity in factor 2 high organoids compared to factor 2 low organoid lines.

