---
title: "Differential gene expression across organoid phenotypes"
author: "Benedikt Rauscher"
date: "11/15/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

# Dependencies

```{r}
library(SummarizedExperiment)
library(DESeq2) #optional
library(glmnet)
library(pheatmap)
library(reshape2)
library(tidyverse)
library(cowplot)
library(here)
```

Define function namespace.

```{r}
filter <- dplyr::filter
select <- dplyr::select
rename <- dplyr::rename
slice <- dplyr::slice
count <- dplyr::count
```

Set parameters for plotting.

```{r}
theme_set(theme_cowplot())
```

# Data loading

We load the gene expression data from an Rdata object from file.

```{r}
load(here('data/processed/expression/promise_expr.rda'))

organoid_morphology <- read_delim(here::here("references/imaging/visual_classification_organoids.csv"), ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::select(line = organoid, morphology = visual_inspection_v2)

## annotate phenotype group
solid <- c('D004', 'D007', 'D010', 'D019', 'D020', 
           'D022', 'D046', 'D054', 'D055')
cystic <- c('D013', 'D018', 'D021', 'D027', 'D030')

## long data frame
promise_long <- assays(promise_expr)$expr %>% 
  as_tibble(rownames = 'probe') %>% 
  pivot_longer(values_to = 'expr', names_to = 'id', -probe) %>%
  left_join(as_tibble(rowData(promise_expr), rownames = 'probe')) %>%
  inner_join(as_tibble(colData(promise_expr), rownames = 'id')) %>%
  select(-chip_name)

## adding phenotype information
promise_long <- promise_long %>% 
  mutate(phenotype = ifelse(line %in% solid, 'solid', 
                     ifelse(line %in% cystic, 'cystic', 'other'))) %>% 
  filter(phenotype != 'other')
```

# Quality control

We perform a clustering to check if the samples are grouped by line of origin.

```{r}
## make expr matrix
expr_mat <- assays(promise_expr)$expr

## generate heatmap annotation
anno <- promise_long %>% 
  distinct(line, rep, phenotype) %>% 
  unite(id, line, rep) %>% 
  mutate_all(as.factor) %>%
  as.data.frame() %>% column_to_rownames('id')

## visualize distance matrix
pheatmap(as.matrix(dist(t(expr_mat))),
         annotation_row = anno)

## visualize most variable probes
pheatmap(expr_mat[order(apply(expr_mat, 1, sd), decreasing = T)[1:500],], 
         annotation_col = anno)
```

'D021_2017-05-19' looks like a technical outlier so we exclude D021T01 from the data. We do not see a clear separation of cystic and solid organoid lines based on the clustering.

D054T01 and D055T01 have been shown to be cross-contaminated, most likely from D046T01. We remove the samples from the dataset.

```{r}
## exclude outlier
promise_long <- promise_long %>% filter(!line %in% c('D054', 'D055', 'D021'))
```

# Unsupervised analysis

I perform a principal component analysis of individual lines at the gene level in order to develop insights into which factors drive variation in the gene expression data.

```{r}
## select most highly expressed probe to represent each gene
select_probes <- promise_long %>% group_by(symbol, probe) %>% 
  summarise(avg_probe = mean(expr)) %>% ungroup() %>%
  group_by(symbol) %>% top_n(1, avg_probe) %>% ungroup() %>% pull(probe)

## summarize replicates
gene_expr <- promise_long %>% 
  # group_by(line, symbol, probe, phenotype) %>%
  # summarise(expr = mean(expr)) %>% ungroup() %>%
  filter(probe %in% select_probes)

## get to 1000 variable genes, do PCA
gene_expr_mat <- gene_expr %>% acast(symbol ~ line + rep, 
                                     value.var = 'expr')
pca <- prcomp(t(gene_expr_mat[order(apply(gene_expr_mat, 1, sd), decreasing = T)[1:1000],]))

## annotate and plot pca results
pca_res <- as_tibble(pca$x, rownames = 'id') %>% select(id, PC1:PC5) %>% 
  inner_join(distinct(gene_expr %>% distinct(id, phenotype))) %>% 
  inner_join(gene_expr %>% filter(symbol == 'MIR483') %>% 
               distinct(id, IGF = expr))

pca_res %>% ggplot(aes(PC1, PC2, color = IGF, shape = phenotype)) +
  geom_point(size = 2) + 
  ggrepel::geom_text_repel(aes(label = id), min.segment.length = 0) + 
  scale_color_gradient(low = '#cccccc', high = '#111111')
```

```{r}
top_genes = select_probes %>% length()
top_genes = top_genes * 0.05
top_genes
```


I plot H19/IGF2 imprinting genes across lines.

```{r}
promise_long %>% mutate(symbol = ifelse(probe == '202410_x_at', 'IGF2', symbol)) %>%
  filter(symbol %in% c('H19', 'MIR483', 'IGF2')) %>%   
  ggplot(aes(line, expr, fill = phenotype)) + 
  stat_summary(fun = 'mean', geom = 'bar') + 
  geom_point() + 
  facet_wrap(~ symbol + probe, ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

promise_long %>% filter(symbol %in% c('PEG10') | (probe == '214218_s_at')) %>%   
  ggplot(aes(line, expr, fill = phenotype)) + 
  stat_summary(fun = 'mean', geom = 'bar') + 
  geom_point() + 
  facet_wrap(~ symbol + probe, ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Expression markers of drug response

Do any of the gene expression probes predict whether an organoid line will respond to a drug treatment? I fit a penalized regression model with the AUROC curve from the SVM analysis as dependent and the gene expression measurements as independent variables. I use a LASSO appraoch to cut down on the number of coefficients since my sample sizes are very low.

```{r}
## load auroc from file
data('aucroc', package = 'SCOPEAnalysis')

## exclude lknes 21, 54, 55
aucroc <- aucroc %>% filter(!line %in% c('D055T01', 'D021T01', 'D020T02', 'D021T01'))

## matrix with independnt variables (gene expression), merge reps
x_lasso <- promise_long %>% 
  group_by(line, symbol, probe, phenotype) %>%
  summarise(expr = mean(expr)) %>% ungroup() %>%
  filter(probe %in% select_probes) %>%
  mutate(line = paste0(line, 'T01')) %>%
  acast(line ~ symbol, value.var = 'expr')
## center by subtracting means and retain genes with high var.
x_lasso <- apply(x_lasso, 2, function(x) x - mean(x))
x_lasso <- x_lasso[,order(apply(x_lasso, 2, var), decreasing=T)[1:1000]]

drug_expr <- aucroc %>% group_by(drug) %>%
  filter(!any(is.na(auroc))) %>%
  group_modify(~{
    ## fit model
    stopifnot(identical(rownames(x_lasso), .x$line))
    fit <- cv.glmnet(
      x = x_lasso,
      y = .x$auroc,
      family = 'gaussian',
      # alpha = 0.5, 
      grouped = F
    )
    ## extract non-zero coefficients
    as_tibble(as.matrix(coef(fit, s = fit$lambda.min)),
              rownames = 'symbol') %>%
      `colnames<-`(c('symbol', 'coef')) %>%
      filter(coef > 0)
  }) %>% ungroup()

## check best coefficients
drug_expr <- drug_expr %>% filter(symbol != '(Intercept)') %>% 
  arrange(desc(coef))
```

Check if there are correlations in the gene expression data for some of the top coefficients.

```{r}
plot(filter(aucroc, drug == 'AZ 960')$auroc, x_lasso[,'KIT'])

plot(filter(aucroc, drug == 'AZD1080')$auroc, x_lasso[,'TUBB6'])

plot(filter(aucroc, drug == 'WYE-125132 (WYE-132)')$auroc, x_lasso[,'LEF1'])

plot(filter(aucroc, drug == 'WYE-125132 (WYE-132)')$auroc, x_lasso[,'H19'])
```

I load drug target annotations from file. Do the same drugs have the same targets?

```{r, eval = FALSE}
## load drug annotation
drug_anno <- readxl::read_excel('data/Compound_Annotation_Libraries_New.xlsx') %>% distinct(drug = `Compound name`, target = `Primary Target`)

## annotate model coefficients
drug_expr <- drug_expr %>% left_join(drug_anno)
```

Generate network diagram.

```{r, eval = FALSE}
library(tidygraph)
library(ggraph)

## correlation between gene expr as links
genes <- drug_expr %>% pull(symbol)
gene_cor <- cor(x_lasso[,colnames(x_lasso) %in% genes])
diag(gene_cor) <- NA
gene_cor <- gene_cor %>% as_tibble(rownames = 'source') %>%
  pivot_longer(names_to = 'target', values_to = 'val', -source) %>%
  filter(val > 0.85) %>% 
  mutate(edge_type = 'gene_cor')

## lasso coefficients
drug_assoc <- drug_expr %>% 
  select(source = drug, target = symbol, val = coef) %>%
  mutate(edge_type = 'drug_assoc')

## combine and make graph
g <- gene_cor %>% bind_rows(drug_assoc) %>% 
  as_tbl_graph() %>% 
  activate(nodes) %>%
  arrange(name) %>%
  mutate(type = ifelse(name %in% genes, 'gene', 'drug'))

## drug target annotation for network
targ_anno <- drug_anno %>% rename(name = drug) %>% 
  bind_rows(distinct(tibble(name = genes, target = 'none'))) %>% 
  filter(!is.na(name), name %in% as_tibble(g)$name) %>%  
  arrange(name)

g <- g %>% activate(nodes) %>% mutate(target = targ_anno$target)

## visualize
g %>% ggraph(layout = 'kk') + 
  geom_edge_link(aes(color = edge_type)) + 
  geom_node_point(size = 2, aes(color = target)) +
  geom_node_text(aes(label = name), colour = 'black', vjust = 0.4) 
```



# Session info

```{r, results = 'show'}
sessionInfo()
```

